---
title: "Solar cell data analysis"
output: github_document
---
# Analysis of solar cell production

Over the last 6 years I recorded the daily production of out roof mounted solar panels. The panels have a peak production of 10kw/p

## Setup
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load solarzellen data
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(ISLR)
library(MASS)
library(car)
source("summary_se.R")
source("solar_auswertung.R")
```

## Read data
Data were recorded daily and put in a spreadsheet. For every year a seperate spreadsheet was used. Data were then exported in csv-format, read into R and compined into one tible.


```{r read_data, echo=FALSE, message=FALSE}
dir_name = "solarzellen_data"
solarcell_files_names = list.files(path = dir_name, pattern="/*.csv")
solarcell_files_names<- paste(dir_name, solarcell_files_names, sep = "/")
solarcell_tbl <- data_frame("datum"=character(0), "zaehlerstand_morgens"= numeric(0), "zaehlerstand_abends" = numeric(0),"produktion"= numeric(0),"eigen_verbrauch"= numeric(0),"eingespeist"= numeric(0))

solarcell_tbl = lapply(solarcell_files_names, read_csv2) %>% bind_rows()
#using lubridate to make the 'Datum' column more usefull
solarcell_tbl$Datum <- dmy(solarcell_tbl$Datum)
solarcell_tbl$YEAR <- year(solarcell_tbl$Datum)
solarcell_tbl$MONTH <- month(solarcell_tbl$Datum)
solarcell_tbl$DAY <- day(solarcell_tbl$Datum)

```
## First processing
### Add standard deviation and mean
```{r basic_stats}
solarcell_tbl_avg <- summarySE(solarcell_tbl, measurevar="Produktion", groupvars=c("YEAR","MONTH"))
#summary over for months
solarcell_tbl_avg %>% 
  group_by( MONTH) %>% 
  summarise_at(vars(Produktion), funs(mean,sd))

#summary over for year
solarcell_tbl_avg %>% 
  group_by(YEAR) %>% 
  summarise_at(vars(Produktion), funs(mean,sd))
```

### Add classification 
Any daily production exceeding 40 kW is considered a fine day, between 20 and 40 kW middling, and anything below is considered a bad day.

```{r weather_classification_1}
solarcell_tbl$rating <- sapply(solarcell_tbl$Produktion, wetter)
```

```{r prepare_fit}
y <- solarcell_tbl$Produktion
t <- 1:nrow(solarcell_tbl)
x <- solarcell_tbl$Datum


# sinus fit 
#1
ssp <- spectrum(y) 
per <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
reslm <- lm(y ~ sin(2*pi/per*t)+cos(2*pi/per*t))
summary(reslm)

rg <- diff(range(y))
plot(y~t,ylim=c(min(y)-0.1*rg,max(y)+0.1*rg))
lines(fitted(reslm)~t,col=4,lty=1)   # dashed blue line is sin fit

# including 2nd harmonic really improves the fit
reslm2 <- lm(y ~ sin(2*pi/per*t)+cos(2*pi/per*t)+sin(4*pi/per*t)+cos(4*pi/per*t))
summary(reslm2)
lines(fitted(reslm2)~t,col=3) 

#2 fft
raw.fft = fft(y)
```
## First Plot

###Scatter plot with sinus fit
```{r  scatter_plot, echo=FALSE}
ggplot(data = solarcell_tbl, aes(x = Datum, y = Produktion)) + 
  geom_point() + 
  geom_smooth(se=FALSE, method="lm", formula = y ~ sin(2*pi/per*x)+cos(2*pi/per*x))
```
### Add classification variables for year, month and day
```{r}
solarcell_tbl$YEAR <- year(solarcell_tbl$Datum)
solarcell_tbl$MONTH <- month(solarcell_tbl$Datum)
solarcell_tbl$DAY <- day(solarcell_tbl$Datum)
```

###Scatter plot by year
```{r scatter_plot_year, echo=FALSE }

ggplot(data = solarcell_tbl, aes(x = MONTH, y = Produktion )) + 
  geom_point() +
  facet_wrap(~YEAR) +
  geom_smooth(se=FALSE, method="lm", formula = y ~ sin(2*pi/per*x)+cos(2*pi/per*x)+sin(4*pi/per*x)+cos(4*pi/per*x))


```
### Monthly averages plot

```{r scatter_plot_month, echo=FALSE }
ggplot(data = solarcell_tbl_avg, aes(x = MONTH, y = Produktion )) + 
  geom_bar(aes(fill=YEAR), position=position_dodge(.9), stat="identity") +
  facet_wrap(~YEAR) +
  geom_errorbar(aes(ymin=Produktion-se, ymax=Produktion+se),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9))
```

### Plot months in a row
```{r monthly_solar_cell_production}
xx <-1:nrow(solarcell_tbl_avg)

ggplot(data = solarcell_tbl_avg, aes(x=xx, y = Produktion )) + geom_bar( stat="identity", width = 0.5)
  
```

## And now for the weather
How many days with fine weather do we have here in Stuttgart

```{r weather_classification_2, echo=FALSE}
ggplot(data = solarcell_tbl, aes(rating)) + geom_histogram(stat="count") + facet_wrap(~YEAR)

```

## Atmospheric conditions
The question was how many days the weather here stays constant, i.e. how many days of fine weather do we have in series. As it turns out, we have a weather here, which changes quite often. Suprise ;(

```{r atmospheric_conditions }
atmospheric_conditions_df <- atm_condition_fn(solarcell_tbl, "rating")
atmospheric_conditions_df$atm_condition <- factor(atmospheric_conditions_df$atm_condition)
ggplot(data = atmospheric_conditions_df, aes(x=duration, fill = atm_condition)) + geom_histogram(binwidth=.5, position = "dodge") +
  scale_x_continuous(breaks = round(seq(min(atmospheric_conditions_df$duration), max(atmospheric_conditions_df$duration), by = 1),1)) 


```
```{r atmospheric_conditions_facet }

ggplot(data = atmospheric_conditions_df, aes(x=duration)) + geom_histogram(binwidth=.5, position = "identity") +
  facet_wrap(~atm_condition)
  scale_x_continuous(breaks = round(seq(min(atmospheric_conditions_df$duration), max(atmospheric_conditions_df$duration), by = 1),1)) 


```
## Read DWD Weather Data from Stuttgart Airport
The data can be obtained from: "https://www.dwd.de/DE/leistungen/weste/westexl/weste_xl.html". You need to register, which can be done at no charge. I have chosen 'sun shine duration', 'top windspeed' and 'mean temperature'. The data are on a per day basis.

```{r read_weather_data}
weather_data <- read_csv2(file = "weste_product_2018_03_10__09_21_28/weste_product_2018_03_10__09_21_27.csv")
weather_data$Datum <- parse_date_time(weather_data$Datum, "ymd")
str(weather_data)
```
### tidy the waether_data
as stated in "http://tidyr.tidyverse.org"
The goal of tidyr is to help you create tidy data. Tidy data is data where:

- Each variable is in a column.
- Each observation is a row.
- Each value is a cell.
But in the weather_data from 'weste-xl' in the column 'Element' are 3 levels, representing 3 different measurement variables (sun shine duraration wind speed and temperature)

```{r}
tmp <- weather_data[c("Datum", "Element", "Wert")]
temp2 <- tidyr::spread(tmp, key = Element, value = Wert)
temp2$Datum <- ymd(temp2$Datum)
solarcell_tbl <- inner_join(solarcell_tbl, temp2)
```
## Predict daily power production from sun shine duration

```{r linear_fit_produktion}
fit_sonne_prod <- lm(solarcell_tbl$Produktion ~ solarcell_tbl$Sonnenscheindauer)
prd <- predict(fit_sonne_prod, interval="prediction")
solarcell_tbl_fit <- cbind(solarcell_tbl, prd)

ggplot(data = solarcell_tbl_fit, aes(x = Sonnenscheindauer, y = Produktion)) + 
  geom_point() +
  geom_line(aes(y=lwr), color = "red", linetype = "dashed") +
  geom_line(aes(y=upr), color = "red", linetype = "dashed") +
  geom_smooth(method=lm, level=0.99, se=TRUE, color="darkred", fill="blue") 
#summary(fit_sonne_prod)fit_sonne_prod$coefficients
summary(fit_sonne_prod)
```
```{r polynomial_fit_produktion}
fit_sonne_prod4 <- lm(solarcell_tbl$Produktion ~ poly(solarcell_tbl$Sonnenscheindauer, 4, raw =T), data = solarcell_tbl)
prd4 <- predict(fit_sonne_prod4, interval="prediction")
solarcell_tbl_fit4 <- cbind(solarcell_tbl, prd)
ggplot(data = solarcell_tbl_fit4, aes(x = Sonnenscheindauer, y = Produktion)) + 
  geom_point() + 
  geom_smooth(se=TRUE, method="lm", formula = y ~ poly(x, 4, raw =T), color="darkred", fill="blue", level =0.99)

```


### Use wind and temperatur to get better predictions
###using a linear model
```{r fit_sonne_prod_multi}
fit_sonne_prod_multi <- lm(solarcell_tbl$Produktion ~ solarcell_tbl$Sonnenscheindauer +  solarcell_tbl$`Lufttemperatur Tagesmittel` + solarcell_tbl$Windspitze)
summary(fit_sonne_prod_multi)

vif(fit_sonne_prod_multi)

```

You can see that Wind has no discrenible effect on the production of the solar cells, but temperature clearly has. The variance inflation factors calculated by vif() show that variable are uncorrelated from each other. To see if there is a wind chilling effect we would have to compare the power production on days with high sun duration and low wind with days with high sun duration and low wind speeds.
### using a polynomial model
```{r polym}
fit_sonne_prod_polym <- lm(solarcell_tbl$Produktion ~ polym( solarcell_tbl$Sonnenscheindauer, solarcell_tbl$`Lufttemperatur Tagesmittel`,solarcell_tbl$Windspitze, degree = 2, raw=TRUE))
summary(fit_sonne_prod_polym)
```
Judging from the F statistics using a polynomial model did not improve on the linear model.