---
title: "Solar cell data analysis"
output: github_document
---
# Analysis of solar cell production

Over the last 6 years I recorded the daily production of out roof mounted solar panels. The panels have a peak production of 10kw/p

## Setup
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load solarzellen data
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
source("summary_se.R")
source("solar_auswertung.R")
```

## Read data
Data were recorded daily and put in a spreadsheet. For every year a seperate spreadsheet was used. Data were then exported in csv-format, read into R and compined into one tible.


```{r read_data, echo=FALSE, message=FALSE}
dir_name = "solarzellen_data"
solarcell_files_names = list.files(path = dir_name, pattern="/*.csv")
solarcell_files_names<- paste(dir_name, solarcell_files_names, sep = "/")
solarcell_tbl <- data_frame("datum"=character(0), "zaehlerstand_morgens"= numeric(0), "zaehlerstand_abends" = numeric(0),"produktion"= numeric(0),"eigen_verbrauch"= numeric(0),"eingespeist"= numeric(0))

solarcell_tbl = lapply(solarcell_files_names, read_csv2) %>% bind_rows()
#using lubridate to make the 'Datum' column more usefull
solarcell_tbl$Datum <- dmy(solarcell_tbl$Datum)
solarcell_tbl$YEAR <- year(solarcell_tbl$Datum)
solarcell_tbl$MONTH <- month(solarcell_tbl$Datum)
solarcell_tbl$DAY <- day(solarcell_tbl$Datum)

```
## First processing
### Add standard deviation and mean
```{r basic_stats}
solarcell_tbl_avg <- summarySE(solarcell_tbl, measurevar="Produktion", groupvars=c("YEAR","MONTH"))
#summary over for months
solarcell_tbl_avg %>% 
  group_by( MONTH) %>% 
  summarise_at(vars(Produktion), funs(mean,sd))

#summary over for year
solarcell_tbl_avg %>% 
  group_by(YEAR) %>% 
  summarise_at(vars(Produktion), funs(mean,sd))
```

### Add classification 
Any daily production exceeding 40 kW is considered a fine day, between 20 and 40 kW middling, and anything below is considered a bad day.

```{r weather_classification}
solarcell_tbl$rating <- sapply(solarcell_tbl$Produktion, wetter)
```

```{r prepare_fit}
y <- solarcell_tbl$Produktion
t <- 1:nrow(solarcell_tbl)
x <- solarcell_tbl$Datum


# sinus fit 
#1
ssp <- spectrum(y) 
per <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
reslm <- lm(y ~ sin(2*pi/per*t)+cos(2*pi/per*t))
summary(reslm)

rg <- diff(range(y))
plot(y~t,ylim=c(min(y)-0.1*rg,max(y)+0.1*rg))
lines(fitted(reslm)~t,col=4,lty=1)   # dashed blue line is sin fit

# including 2nd harmonic really improves the fit
reslm2 <- lm(y ~ sin(2*pi/per*t)+cos(2*pi/per*t)+sin(4*pi/per*t)+cos(4*pi/per*t))
summary(reslm2)
lines(fitted(reslm2)~t,col=3) 

#2 fft
raw.fft = fft(y)
```
## First Plot

###Scatter plot with sinus fit
```{r  scatter_plot, echo=FALSE}
ggplot(data = solarcell_tbl, aes(x = Datum, y = Produktion)) + 
  geom_point() + 
  geom_smooth(se=FALSE, method="lm", formula = y ~ sin(2*pi/per*x)+cos(2*pi/per*x))
```
### Add classification variables for year, month and day
```{r}
solarcell_tbl$YEAR <- year(solarcell_tbl$Datum)
solarcell_tbl$MONTH <- month(solarcell_tbl$Datum)
solarcell_tbl$DAY <- day(solarcell_tbl$Datum)
```

###Scatter plot by year
```{r scatter_plot_year, echo=FALSE }

ggplot(data = solarcell_tbl, aes(x = MONTH, y = Produktion )) + 
  geom_point() +
  facet_wrap(~YEAR) +
  geom_smooth(se=FALSE, method="lm", formula = y ~ sin(2*pi/per*x)+cos(2*pi/per*x)+sin(4*pi/per*x)+cos(4*pi/per*x))


```
### Monthly averages plot

```{r scatter_plot_month, echo=FALSE }
ggplot(data = solarcell_tbl_avg, aes(x = MONTH, y = Produktion )) + 
  geom_bar(aes(fill=YEAR), position=position_dodge(.9), stat="identity") +
  facet_wrap(~YEAR) +
  geom_errorbar(aes(ymin=Produktion-se, ymax=Produktion+se),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9))
```

### Plot months in a row
```{r monthly_solar_cell_production}
xx <-1:nrow(solarcell_tbl_avg)

ggplot(data = solarcell_tbl_avg, aes(x=xx, y = Produktion )) + geom_bar( stat="identity", width = 0.5)
  
```

## And now for the weather
How many days with fine weather do we have here in Stuttgart

```{r weather_classification, echo=FALSE}
ggplot(data = solarcell_tbl, aes(rating)) + geom_histogram(stat="count") + facet_wrap(~YEAR)

```

## Atmospheric conditions

```{r atmospheric_conditions }
atmospheric_conditions_df <- atm_condition_fn(solarcell_tbl, "rating")
atmospheric_conditions_df$atm_condition <- factor(atmospheric_conditions_df$atm_condition)
ggplot(data = atmospheric_conditions_df, aes(x=duration, fill = atm_condition)) + geom_histogram(binwidth=.5, alpha=.5, position = "dodge")


```

